Skew-symmetric entropy stable discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes

### 2018-08-13 docs/skewform/main.tex
updating with new skew form writing

%\documentclass[preprint,10pt]{elsarticle}
%\documentclass[preprint,10pt]{article}
\documentclass[review]{siamart0216}
%\documentclass{siamart0216}

%\usepackage{fullpage}
%\usepackage[colorlinks=true]{hyperref}

\usepackage{amsmath,amssymb,amsfonts}
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{definition}{Definition}
%\theoremstyle{lemma}
%\newtheorem{lemma}{Lemma}
%\theoremstyle{corollary}
%\newtheorem{corollary}{Corollary}
%\theoremstyle{theorem}
%\newtheorem{theorem}{Theorem}
\newtheorem*{remark}{Remark}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}

@@

\renewcommand{\hat}[1]{\widehat{#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

@@

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\widehat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

@@

%\begin{frontmatter}
%\title{Entropy stable skew-symmetric discontinuous Galerkin formulations with applications to hybrid and non-conforming meshes. }
%
%\author[rice]{Jesse Chan\corref{cor1}}
%\ead{Jesse.Chan@caam.rice.edu}
%\address[rice]{Department of Computational and Applied Mathematics, Rice University, 6100 Main St, Houston, TX, 77005}
%
%\begin{abstract}
%Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  
%\end{abstract}
%\end{frontmatter}

%\tableofcontents

@@

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule.  In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely}.  These methods have also been extended to staggered grid methods \cite{parsani2016entropy} and non-conforming meshes \cite{friedrich2017entropy}.  Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.

@@

\section{Approximation spaces}

We define an approximation space using degree $N$ polynomials on the reference element.  In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).  The natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.

On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
\[
P^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
\]
In contrast, the natural polynomial space on a $d$-dimensional tensor product element is 
\[
Q^N\LRp{\widehat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \widehat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
\]
We take $V^N$ to denote the natural approximation space on a given element.  Furthermore, we denote the dimension of $V^N$  as $N_p = {\rm dim}\LRp{V^N\LRp{\widehat{D}}}$.

We also define trace spaces for each reference element.  Let $\hat{f}$ be a face of the reference element $\hat{D}$.  The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$
\[
V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
\]
For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.  We denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.

We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
\[
V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
\]

\section{$L^2$ projection and lifting matrices and a decoupled SBP operator}

The schemes in \cite{chan2017discretely} begin on the reference element $\hat{D}$ by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$.  It is assumed that volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ are provided for $\hat{D}$ and the surface $\partial \hat{D}$ of $\hat{D}$, respectively.  We make the following assumption on the volume quadrature rule:
\begin{assumption}
The volume quadrature rule integrates exactly all polynomials of degree $(2N-1)$.
\label{ass:quad}
\end{assumption}
In addition to being related to the positive definiteness of the mass matrix, Assumption~\ref{ass:quad} ensures that, for any $u,v \in V^N$, the following integral is exactly approximated using quadrature
\[
\int_{\hat{D}} \pd{u(\hat{\bm{x}})}{\hat{x}_i} v(\hat{\bm{x}})\diff{\hat{\bm{x}}} = \sum_{k=1}^{N_q} w_k \pd{u(\hat{\bm{x}}_k)}{\hat{x}_i} v(\hat{\bm{x}}_k).  
\]
In other words, evaluating the integral of the derivative of $u$ and $v$ using quadrature is exact because the integrand is a polynomial of degree $(2N-1)$.

\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}

@@

Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.  For example, the matrix $\bm{D}_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.  By choosing specific quadrature rules, $\bm{D}_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.  However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.

Let $\bm{n}$ denote the outward unit normal $\bm{n}$ on $\hat{D}$, and let $\bm{n}_i$ denote the vector containing values of the $i$th component at all surface quadrature points.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{align}
\bm{D}^i_N  &= \LRs{
\begin{array}{cc}
\bm{D}^i_q - \frac{1}{2}\bm{V}_q \bm{L}_f {\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\bm{V}_q\bm{L}_f{\rm diag}({\bm{n}}_i)\\
-\frac{1}{2}{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}{\rm diag}({\bm{n}}_i)
\end{array}}.  
\label{eq:DN}
\end{align}
When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that 
\[
\int_{\hat{D}} u(\bm{x})v(\bm{x})  = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}n_i, \qquad \forall v\in V^N\LRp{\hat{D}}.
\]
If the surface quadrature rule is exact for polynomials of degree $2N$, then the decoupled SBP operator $\bm{D}^i_N$ also satisfies a summation by parts property.  
\begin{gather}
\bm{Q}^i_N = \bm{W}_N \bm{D}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:dsbp}\\ \bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\nonumber
\end{gather}
By noting that $\bm{W}\bm{V}_q \bm{L}_q = \bm{P}_q^T\bm{V}_f^T\bm{W}_f$, the entries of $\bm{Q}^i_N$ can be re-written as
\begin{align}
\bm{Q}_i &= \bm{W}\bm{V}_q\bm{D}^i\bm{P}_q,\\
\bm{Q}^i_N  &= \LRs{
\begin{array}{cc}
\bm{Q}_i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T  \bm{W}_f{\rm diag}({\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q}^T\\
-\frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)\bm{V}_f\bm{P}_q & \frac{1}{2}\bm{W}_f{\rm diag}({\bm{n}}_i)
\end{array}}.
\label{eq:QN}
\end{align}
from which the SBP property (\ref{eq:dsbp}) is more apparent.

%\begin{lemma}[Lemma 1 in \cite{chan2017discretely}]
%Let $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$.  Then, for volume quadratures which are exact for degree $2N-1$ polynomials, $\bm{D}^i_q$ satisfies the SBP property with respect to $\bm{W}$ 
%\[
%\bm{W}\bm{D}^i_q = \LRp{\bm{V}_f\bm{P}_q}^T \bm{W}_f\diag{\hat{\bm{n}}} \bm{V}_f\bm{P}_q - \LRp{\bm{W}\bm{D}^i_q}^T.  
%\]
%\label{lemma:gsbp}
%\end{lemma}

@@

\subsection{Loss of the SBP property for hybrid and non-conforming meshes}

While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.  This difficulty is a result of the choices of surface quadrature which naturally arise for these settings.

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in an SBP property and are not entropy stable.  On the left, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.  On the right, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral.  }
\label{fig:hybrid}
\end{figure}

\begin{figure}
\centering
\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
\hspace{2em}
\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
\label{fig:hybrid3d}
\end{figure}

\begin{figure}[!h]
\centering
\begingroup
\captionsetup[subfigure]{width=.45\textwidth}
\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
\hspace{2em}
\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.  The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
\endgroup
%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.  Each dashed red line indicates a dyadic flux computation required between two nodes. } %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.  The new approach results in a simpler communication pattern between elements. }
\label{fig:noncon}
\end{figure}

@@

It can be shown that the SBP property holds for $\bm{D}^i_N$ if the surface quadrature is exact for polynomials of degree $2N$ \cite{chan2017discretely}.  The SBP property is lost if the surface quadrature is only exact for polynomial integrands of degree less than $2N$.  However, one can still show that a variational version of the SBP property is satisfied.

>>>>

\documentclass[review]{siamart0216}

\usepackage{amsmath,amssymb,amsfonts}
\newtheorem*{remark}{Remark}
\theoremstyle{assumption}
\newtheorem{assumption}{Assumption}

@@

\renewcommand{\hat}[1]{\hat{#1}}
\renewcommand{\topfraction}{0.85}
\renewcommand{\textfraction}{0.1}
\renewcommand{\floatpagefraction}{0.75}

@@

\newcommand{\Oh}{{\Omega_h}}
\renewcommand{\L}{L^2\LRp{\Omega}}
\newcommand{\LK}{L^2\LRp{D^k}}
\newcommand{\LdK}{L^2\LRp{\partial D^k}}
\newcommand{\Dhat}{\hat{D}}
\newcommand{\Lhat}{L^2\LRp{\Dhat}}

\renewcommand{\hat}{\widehat}

@@

\begin{abstract}
Entropy stable high order methods for nonlinear conservation laws satisfy an inherent discrete entropy inequality.  The construction of such schemes has relied on the use of carefully chosen collocation points \cite{gassner2013skew, fisher2013high, carpenter2014entropy, chan2018efficient} or volume and surface quadrature rules \cite{chan2017discretely, chan2018discretely} to produce operators which satisfy a summation-by-parts (SBP) property.  In this work, we show how to generalize the construction of semi-discretely entropy stable schemes to accommodate more general pairs of volume and surface quadratures under which an SBP property does not hold.  We conclude by discussing the applicability of these new operators for entropy stable schemes on hybrid and non-conforming meshes.  
\end{abstract}

@@

These issues have motivated the introduction of high order \textit{entropy stable} discretizations, which satisfy a semi-discrete entropy inequality while maintaining high order accuracy in smooth regions.  Proofs of continuous entropy inequalities rely on the chain rule.  In contrast, these discrete entropy inequalities account for the loss of the chain rule due to effects such as quadrature errors, which are incurred when applying polynomially exact quadrature rules to nonlinear and rational integrands within the DG formulation.  These schemes were first introduced as high order collocation methods on tensor product elements in \cite{fisher2013high, carpenter2014entropy, gassner2016split, gassner2017br1}, and were extended to simplicial elements in \cite{crean2017high, chen2017entropy, crean2018entropy, chan2017discretely, chan2018discretely}.  These methods have also been extended to staggered grid \cite{parsani2016entropy}, generalized SBP and Gauss nodes \cite{chan2018efficient} and non-conforming meshes \cite{friedrich2017entropy}.  Entropy stable boundary conditions have also been determined for the compressible Euler and Navier-Stokes equations \cite{parsani2015entropy, svard2018entropy}.

\note{Finish: add Section description.  Talk about situations where the SBP or decoupled SBP property doesn't hold: triangles with reduced surface quadrature, GLL quads with GQ face quadratures, and hybrid couplings.}

@@

\section{Polynomial approximation spaces}

In this work, we consider either simplicial reference elements (triangles and tetrahedra) or tensor product reference elements (quadrilaterals and hexahedra).  We define an approximation space using degree $N$ polynomials on the reference element; however, the natural polynomial approximation space differs depending on the element type \cite{chan2015gpu}.

On a $d$-dimensional reference simplex, the natural polynomial space are total degree $N$ polynomials 
\[
P^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq \sum_{k=1}^d i_k \leq N}.
\]
In contrast, the natural polynomial space on a $d$-dimensional tensor product element is the space of maximum degree $N$ polynomials
\[
Q^N\LRp{\hat{D}} = \LRc{\hat{x}_1^{i_1}\ldots\hat{x}_d^{i_d}, \quad \hat{\bm{x}} \in \hat{D}, \quad 0\leq i_k \leq N, \quad k = 1,\ldots, d}.
\]
We denote the natural approximation space on a given reference element $\hat{D}$ by $V^N$.  Furthermore, we denote the dimension of $V^N$ as $N_p = {\rm dim}\LRp{V^N\LRp{\hat{D}}}$.

We also define trace spaces for each reference element.  Let $\hat{f}$ be a face of the reference element $\hat{D}$.  The trace space $V^N \LRp{\hat{f}}$ is defined as the restrictions of functions in $V^N\LRp{\hat{D}}$ to $\hat{f}$, and denote the dimension of the trace space as ${\rm dim}\LRp{V^N\LRp{{\hat{f}}}} = N^f_p$.  
\[
V^N \LRp{\hat{f}} = \LRc{ \left.u\right|_{\hat{f}}, \quad u \in V^N\LRp{\hat{D}}}, \qquad \hat{f}\in \partial\hat{D}.
\]
For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.  On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.

%For example, the trace space for degree $N$ polynomials on a quadrilateral face $\hat{f}$ of the bi-unit hexahedral element $[-1,1]^3$ is $Q^N\LRp{\hat{f}}$, while the trace space for degree $N$ polynomials on a triangular face $f$ of the tetrahedron is $P^N\LRp{\hat{f}}$.

%We similarly define the trace space for the surface $\partial \hat{D}$ of $\hat{D}$
%\[
%V^N \LRp{\partial \hat{D}} = \LRc{ \left.u\right|_{\partial \hat{D}}, \quad u \in V^N\LRp{\hat{D}}}.
%\]
%For example, on a $d$-dimensional simplex, $V^N \LRp{\partial \hat{D}}$ consists of total degree $N$ polynomials on simplices of dimension $(d-1)$.  On a $d$-dimensional tensor product element, $V^N \LRp{\partial \hat{D}}$ consists of maximum degree $N$ polynomials on a tensor product element of dimension $(d-1)$.

\section{Quadrature-based matrices and decoupled SBP operators}

\note{TODO: Standardize hat notation for operators/normals: can refer to operators on reference element and do $k$ superscript for physical elements, or do hats on reference element.}
Let $\hat{D} \subset\mathbb{R}^d$ denote a reference element with surface $\partial \hat{D}$.  
The high order schemes in \cite{chan2017discretely, chan2018discretely} begin by approximating the solution in a degree $N$ polynomial basis $\LRc{\phi_j({\bm{x}})}_{i=1}^{N_p}$ on $\hat{D}$.  These schemes also assume volume and surface quadrature rules $({\bm{x}}_i, w_i)$, $\LRp{{\bm{x}}^f_i,w^f_i}$ on $\hat{D}$.  We will specify the accuracy of each quadrature rule later, and discuss how quadrature accuracy implies specific summation-by-parts properties.

@@

Interpolation, differentiation, and $L^2$ projection matrices can be combined to construct finite difference operators.  For example, the matrix $\bm{D}^i_q = \bm{V}_q\bm{D}^i\bm{P}_q$ maps function values at quadrature points to approximate values of the derivative at quadrature points.  By choosing specific quadrature rules, $\bm{D}^i_q$ recovers high order summation-by-parts finite difference operators in \cite{gassner2013skew, fernandez2014generalized, ranocha2018generalised} and certain operators in \cite{hicken2016multidimensional}.  However, to address difficulties in designing efficient entropy stable interface terms for nonlinear conservation laws, the PI introduced in \cite{chan2017discretely} a new ``decoupled'' summation by parts matrix which builds interface terms directly into the approximation of the derivative.

Let $\hat{\bm{n}}$ denote the scaled outward normal vector $\hat{\bm{n}} = \LRc{\hat{n}_1\hat{J}_f,\ldots,\hat{n}_d\hat{J}_f}$, where $\hat{J}_f$ is the determinant of the Jacobian of the mapping of a face of $\partial \hat{D}$ to a reference face.  Let $\hat{\bm{n}}_i$ denote the vector containing values of the $i$th component $\hat{n}_i\hat{J}f$ at all surface quadrature points, and let $\bm{Q}^i = \bm{W}\bm{D}^i_q$.  The ``decoupled'' summation by parts operator $\bm{D}^i_N$ is defined as the block matrix involving both volume and surface quadratures
\begin{gather}
\bm{D}^i_N = \bm{W}_N^{-1} \bm{Q}^i_N, \qquad \bm{W}_N = \LRp{\begin{array}{cc}\bm{W}&\\ & \bm{W}_f\end{array}},\label{eq:QN}\\
\bm{Q}^i_N  = \LRs{
\begin{array}{cc}
\bm{Q}^i - \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q &  \frac{1}{2}\LRp{\bm{V}_f \bm{P}_q}^T \bm{W}_f {\rm diag}(\hat{\bm{n}}_i)\\
-\frac{1}{2}\bm{W}_f{\rm diag}(\hat{\bm{n}}_i) \bm{V}_f\bm{P}_q & \frac{1}{2} \bm{W}_f{\rm diag}(\hat{\bm{n}}_i)
\end{array}}.  \nonumber
\end{gather}
When combined with projection and lifting matrices, $\bm{D}^i_N$ produces a high order polynomial approximation of $f\pd{g}{x_i}$. 
Let $f, g$ be differentiable functions, and let $\bm{f}_i = f(\bm{x}_i)$, $\bm{g}_i = g(\bm{x}_i)$ denote values of $f,g$ at volume and surface quadrature points.  Then,
\begin{gather}
f\pd{g}{x_i} \approx \LRs{\begin{array}{cc}
\bm{P}_q & \bm{L}_f\end{array}} {\rm diag}\LRp{\bm{f}}\bm{D}^i_N \bm{g}.  
\label{eq:dnapprox}
\end{gather}
The approximation can also be interpreted as solving a variational problem.  Let $n_i$ be the $i$th component of the unit normal on $\hat{D}$.  Then (\ref{eq:dnapprox}) is equivalent to finding $u(\bm{x}) \approx f\pd{g}{x_i}$ such that, for all $v\in V^N\LRp{\hat{D}}$, 
\begin{align*}
\int_{\hat{D}} u(\bm{x})v(\bm{x}) = \int_{\hat{D}}{g\pd{\Pi_Nf}{x_i}v} + \int_{\partial \hat{D}}{(f-\Pi_Nf)\frac{\LRp{gv + \Pi_N(gv)}}{2}}\hat{n}_i\hat{J}_f.
\end{align*}

The matrix $\bm{Q}^i_N$ also satisfies a decoupled summation-by-parts (SBP) property, which is used to prove semi-discrete entropy stability for nonlinear conservation laws.  
\begin{lemma}
Assume that the volume and surface quadrature rules are sufficiently accurate such that the quantities
\[
\int_{\hat{D}} \pd{u}{x_i} v, \qquad \int_{\partial \hat{D}} u v n_i
\]
are integrated exactly for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.  Then, the decoupled SBP operator $\bm{D}^i_N$ satisfies a summation by parts property:
\begin{gather}
\bm{Q}^i_N+\LRp{\bm{Q}^i_N}^T = \LRp{\begin{array}{cc}\bm{0}&\\ & \bm{W}_f {\rm diag}(\bm{n}_i)\end{array}} = \bm{B}_N\label{eq:dsbp}
\end{gather}
%\begin{enumerate}
%\item The mass matrix is positive definite, 
%\item The quadrature integrates exactly $\int_{\hat{D}} \pd{u}{x_i} v$ for all $u,v \in V^N\LRp{\hat{D}}$ and $i = 1,\ldots, d$.
%\end{enumerate}
\label{lemma:dsbp}
\end{lemma}
\begin{proof}
The proof is a straightforward extension of Theorem 1 in \cite{chan2017discretely} to polynomial approximation spaces on non-simplicial elements.  
\end{proof}

The assumptions of Lemma~\ref{lemma:dsbp} are satisfied for sufficiently accurate volume and surface quadratures.  For example, on simplicial elements, (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of total degree $(2N-1)$, and the surface integral is exact for degree $2N$ polynomials on each face.  Tensor product elements require stricter conditions: (\ref{eq:dsbp}) holds if the volume quadrature is exact for polynomial integrands of degree $2N$ in each coordinate, due to the fact that derivatives of $u\in Q^N$ are degree $(N-1)$ with respect to one coordinate and degree $N$ with respect to others.

\begin{remark}
It is worth noting that the assumptions of Lemma~\ref{lemma:dsbp} are sufficient but not necessary conditions.  For example, suppose $\hat{D}$ is a quadrilateral element, and that the volume quadrature is a tensor product of $(N+1)$ point one-dimensional Gauss-Legendre-Lobatto (GLL) rule.  If the surface quadrature also taken to be an $(N+1)$-point GLL rule over each face, then $\bm{Q}^i$ satisfies a traditional SBP property, which can be used to prove the decoupled SBP property for $\bm{Q}^i_N$.  
\end{remark}

@@

While the SBP property has been used to derive entropy stable schemes, the SBP property is difficult or impossible to enforce in certain discretization settings, such as hybrid and non-conforming meshes arises.  This difficulty is a result of the choices of volume and surface quadrature which naturally arise in these settings.  We first illustrate how specific pairings of volume and surface quadratures can result in the loss of the SBP property (\ref{eq:dsbp}).  We then propose a skew-symmetric formulation which is entropy conservative without explicitly requiring operators which satisfy the SBP property.

\subsection{Loss of the SBP property}

In this section, we give examples of specific pairings of volume and surface quadratures under which the decoupled SBP property does not hold (see Figure~\ref{fig:sbploss}).  We consider two dimensional reference elements $\hat{D}$ with spatial coordinates $x,y$.
\begin{figure}
\centering
\subfloat[GLL volume quadrature, Gauss surface quadrature]{\includegraphics[width=.4\textwidth]{figs/gllgauss.png}\label{subfig:gllgq}}
\hspace{2em}
\subfloat[Degree $2N$ volume quadrature, GLL surface quadrature]{\includegraphics[width=.4\textwidth]{figs/trigll.png}\label{subfig:trigll}}
\caption{Volume and surface quadrature pairs which do not satisfy the assumptions of Lemma~\ref{lemma:dsbp}, and thus do not possess the decoupled SBP property (\ref{eq:dsbp}). }
\label{fig:sbploss}
\end{figure}

\paragraph{Quadrilateral elements (Figure~\ref{subfig:gllgq})} We first consider a quadrilateral element $\hat{D}$ with an $(N+1)$ point tensor product GLL volume quadrature and $(N+1)$ point Gauss quadrature on each face.  Let $u,v \in Q^N$ denote two arbitrary degree $N$ polynomials.  The assumptions of Lemma~\ref{lemma:dsbp} are that the volume quadrature exactly integrates $\int_{\hat{D}} \pd{u}{x} v$ and that the surface quadrature exactly integrates $\int_{\partial \hat{D}} u v n_i$ on $\hat{D}$.  Because the $(N+1)$-point Gauss rule is exact for polynomials of degree $2N+1$ and the product $uv \in P^{2N}$ on each face, the surface quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.  However, the 1D GLL rule is only exact for polynomials of degree $(2N-1)$.  The derivative $\pd{u}{x}$ is a polynomial of degree $(N-1)$ in $x$, but is degree $N$ in $y$.  Thus, $\pd{u}{x}v$ is a polynomial of degree $(2N-1)$ in $x$ but degree $2N$ in $y$, and is not integrated exactly by the volume quadrature.

\paragraph{Triangular elements (Figure~\ref{subfig:trigll})} We next consider a triangular element $\hat{D}$, where the volume quadrature is exact for degree $2N$ polynomials \cite{xiao2010quadrature} and an $(N+1)$-point GLL quadrature on each face.  Let $u,v \in P^N$ denote two arbitrary degree $N$ polynomials.  The derivative $\pd{u}{x} \in P^{(N-1})$, and $\pd{u}{x}v \in P^{(2N-1)}$, so the volume quadrature satisfies the assumptions of Lemma~\ref{lemma:dsbp}.  However, because the surface quadrature is exact only degree $(2N-1)$ polynomials and the trace of $uv\in P^{2N}$, the surface quadrature does not satisfy the assumptions of Lemma~\ref{lemma:dsbp}.

\begin{figure}
\centering
\begingroup
\captionsetup[subfigure]{width=.425\textwidth}
\subfloat[Insufficiently accurate surface quadrature on the triangle element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D.png}\label{subfig:hybrid1}}
\endgroup
\hspace{2em}
\subfloat[Incompatible surface quadrature on the quadrilateral element.]{\includegraphics[width=.425\textwidth]{figs/hybrid2D_GQ.png}\label{subfig:hybrid2}}
\caption{Examples of interface couplings which do not result in a decoupled SBP property (\ref{eq:dsbp}).  }
\label{fig:hybrid}
\end{figure}

These specific pairings of volume and surface quadratures appear naturally for hybrid meshes consisting of DG-SEM quadrilateral elements (using GLL volume quadrature) and triangular elements, as shown in Figure~\ref{fig:hybrid}.  In Figure~\ref{subfig:hybrid1}, the surface quadrature is a $(N+1)$ point GLL rule, and results in a loss of the SBP property on the triangle.  In Figure~\ref{subfig:hybrid2}, the surface quadrature is a $(N+1)$ point Gauss-Legendre rule, and results in a loss of the SBP property on the quadrilateral element.

The goal of this work is to construct high order accurate discretizations which preserve entropy conservation for these situations in which the decoupled SBP property (\ref{eq:dsbp}) does not hold.

%\begin{figure}
%\centering
%\subfloat[Hex-pyramid coupling]{\includegraphics[width=.35\textwidth]{figs/hybrid3D.png}}
%\hspace{2em}
%\subfloat[Hex-prism coupling]{\includegraphics[width=.375\textwidth]{figs/hybrid3D_wedge.png}}
%\caption{Illustration of a 3D coupling between a GLL hexahedral element and a pyramid.  The SBP property does not hold on the pyramid due to the use of GLL quadrature on the quadrilateral face. }
%\label{fig:hybrid3d}
%\end{figure}
%
%\begin{figure}[!h]
%\centering
%\begingroup
%\captionsetup[subfigure]{width=.45\textwidth}
%\subfloat[Entropy stable inter-element coupling in \cite{friedrich2017entropy} for a non-conforming interface.]{\raisebox{0em}{\includegraphics[width=.45\textwidth]{figs/nonconSBP.png}}\label{subfig:noncon2}}
%\hspace{2em}
%\subfloat[Entropy stable inter-element coupling in this work for a non-conforming interface.  The dotted black lines denote communication between neighboring elements.]{\raisebox{-0em}{\includegraphics[width=.45\textwidth]{figs/nonconQuad.png}}\label{subfig:noncon1}}
%\endgroup
%%\subfloat[3D hexahedra-pyramid coupling]{\includegraphics[width=.31\textwidth]{figs/hybrid3D.png}}
%\caption{Illustration of two different entropy stable inter-element couplings for $h$ non-conforming meshes.  Each dashed red line indicates a dyadic flux computation required between two nodes. } %The coupling terms in \cite{friedrich2017entropy} requires computing dyadic fluxes between \emph{each} pair of nodes on adjacent interfaces.  The new approach results in a simpler communication pattern between elements. }
%\label{fig:noncon}
%\end{figure}

@@

%The main restriction we wish to overcome

The property (\ref{eq:dsbp}) (which we will refer to as the ``strong'' SBP property) relates the polynomial exactness of specific quadrature rules to algebraic properties of quadrature-based matrices.  We will relax accuracy conditions on these quadrature rules by utilizing not the strong SBP property, but a weaker ``variational'' version of the SBP property.
